Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-07 metrics.tex
Fixing typo
Final submission version

\subsection{Goodput}
We consider the goodput defined as the ratio of the useful delivered
information at the application layer and the total processing time. We focus
on the goodput considering only the coding process, i.e. we assume that
the application data has been properly generated before encoding and
also correctly post-processed after decoding. In this way, we define
the goodput for either an encoder or a decoder as follows:
%
\begin{align} \label{eq:goodput}
R_{proc} = \frac{gB}{T_{proc}} ~ [\mathrm{Bytes/second}]
\end{align}
%
In \eqref{eq:goodput}, $B$ and $g$ are the packet and generation size
as defined previously and both represented the data to be processed. For
goodput measurements, we are concerned in quantifying the processing time
for either encoding or decoding $g$ \ac{l.i.} packets. Thus, $T_{proc}$ is
the processing time required for this processing. In the next subsections, we
define two time bechmarks available in [\textbf{Chres repos here}].
The purpose of the benchmarks is to quantify the processing time for
any of the coding schemes considered in Section \ref{sec:schemes}.

@@

The average power specifies the rate at which energy is
consumed in the \ac{Raspi}. Thus, for a given energy value in the device
battery without any external supplies, this metric permits to infer the amount
of time for which the \ac{Raspi} can operate autonomously before draining
out its battery. For the energy per bit consumption, it indicates how
much energy is expended to effectively transmit or receive 1 bit of bit
considering the encoding or decoding respectively.

>>>>

\subsection{Goodput}
We consider the goodput defined as the ratio of the useful delivered
information at the application layer and the total processing time. We focus
on the goodput considering only the coding process, i.e. we assume that
the application data has been properly generated before encoding and
also correctly post-processed after decoding. In this way, we define
the goodput for either an encoder or a decoder as follows:
%
\begin{align} \label{eq:goodput}
R_{proc} = \frac{gB}{T_{proc}} ~ [\mathrm{Bytes/second}]
\end{align}
%
In \eqref{eq:goodput}, $B$ and $g$ are the packet and generation size
as defined previously and both represent the data to be processed. For
goodput measurements, we are concerned in quantifying the processing time
for either encoding or decoding $g$ \ac{l.i.} packets. Thus, $T_{proc}$ is
the processing time required for this processing. In the next subsections, we
define two time bechmarks available in \cite{soerensen2016sparse}.
The purpose of the benchmarks is to quantify the processing time for
any of the coding schemes considered in Section \ref{sec:schemes}.

@@

The average power specifies the rate at which energy is
consumed in the \ac{Raspi}. Thus, for a given energy value in the device
battery without any external supplies, this metric permits to infer the amount
of time for which the \ac{Raspi} can operate autonomously before draining
out its battery. For the energy per bit consumption, it indicates how
much energy is expended to effectively transmit or receive 1 bit of
information taking into account the encoding or decoding operations
respectively.

### 2016-06-07 raspi_journal.tex
Final submission version

% Authors, for the paper (add full first names)
\Author{N\'estor J. Hern\'andez Marcano $^{1,2,\dagger}$*, Chres W. S{\o}rensen $^{2,\dagger}$, Juan A. Cabrera G. $^{3,4}$, Simon Wunderlich $^{3}$, Daniel E. Lucani $^{2,\dagger}$ and Frank H. P. Fitzek $^{3}$}
% Authors, for metadata in PDF
\AuthorNames{Nestor Hernandez, Chres W. Soerensen, Juan Cabrera, Simon Wunderlich, Daniel E. Lucani and Frank H. P. Fitzek}
% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Steinwurf ApS. Aalborg, Denmark; nestor@steinwurf.com\\
$^{2}$ \quad Aalborg University, Department of Electronic Systems. Aalborg, Denmark; \{nh, cws, del\}@es.aau.dk\\
$^{3}$ \quad Technische Universit\"at Dresden, Deutsche Telekom Chair of Communication Networks. Dresden, Germany; \{juan.cabrera, simon.wunderlich, frank.fitzek\}@tu-dresden.de\\
$^{4}$ \quad SFB 912 -- Collaborative Research Center HAEC. Dresden, Germany}

>>>>

% Authors, for the paper (add full first names)
\Author{Chres W. S{\o}rensen $^{1,\dagger}$, N\'estor J. Hern\'andez Marcano $^{1,2,\dagger}$*, Juan A. Cabrera G. $^{3,4}$, Simon Wunderlich $^{3}$, Daniel E. Lucani $^{1,\dagger}$ and Frank H. P. Fitzek $^{3}$}
% Authors, for metadata in PDF
\AuthorNames{Chres W. Soerensen, Nestor J. Hernandez M., Juan Cabrera, Simon Wunderlich, Daniel E. Lucani and Frank H. P. Fitzek}
% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Aalborg University, Department of Electronic Systems. Aalborg, Denmark; \{nh, cws, del\}@es.aau.dk\\
$^{2}$ \quad Steinwurf ApS. Aalborg, Denmark; nestor@steinwurf.com\\
$^{3}$ \quad Technische Universit\"at Dresden, Deutsche Telekom Chair of Communication Networks. Dresden, Germany; \{juan.cabrera, simon.wunderlich, frank.fitzek\}@tu-dresden.de\\
$^{4}$ \quad SFB 912 -- Collaborative Research Center HAEC. Dresden, Germany}

### 2016-06-07 testbed.tex
Final submission version

%This section will present a testbed of thirty \ac{Raspi}s and a \ac{PC}
The testbed consists of thirty \ac{Raspi}s that will be setup and controlled
by a testbed administrator from a \ac{PC}.
%Twenty \ac{Raspi} 1 model B revision 2 and ten \ac{Raspi} 2 model B V1.1.
The \ac{Raspi}s are each equipped with a memory card and connected to the
university network by wire using their ethernet interface (eth0).

To easily discover the \ac{Raspi}s within the university network,
they have all been assigned a unique name under a common domain.
This eliminates the need to remember their individual \ac{IP} addresses.

A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
We have two types of \ac{Raspi}s in our setup.
Table~\ref{tab:rasp_naming} shows the naming convention.

\begin{table}[ht!]
  \centering
  \begin{tabular}{|c|c|}
    \hline
      \textbf{Hostname} & \textbf{Model} \\ \hline
    Rasp00 - Rasp19 & Raspberry Pi 1 model B rev. 2 \\ \hline
    Rasp20 - Rasp29 & Rasperry Pi 2 model B V1.1 \\ \hline
  \end{tabular}
  \caption{Raspberry Pi naming convention}
  \label{tab:rasp_naming}
\end{table}

@@

The testbed should be usable by students to perform simulations and
measurements as part of their semester projects. It is therefore desired
that the \ac{Raspi}s are configured identically, but still allows the
students to alter the configurations.

When working on a testbed, it is not always desired to make persistent
changes to the devices. We also present how the concept of stacked
filesystems can be used to enable the means of both persistent and
temporary storage.

@@

To get started, we first need to install an \ac{OS} on the
\ac{Raspi}s. We will use Raspbian linux~\cite{raspbian}

We will be using a \ac{PC} running a Debian-based Linux distribution
to setup a customized Raspbian Linux image that can be written to
a memory card and work in any \ac{Raspi} version 1 or 2.
%We will use a desktop that has linux installed to download,
%configure, and write the image to the memory cards.
%We have decided to use a linux desktop machine to create a
%common image of Raspbian Linux that can be used in any of the Raspberry Pis.

@@

Raspbian is made available in two bundles. Raspbian and Raspbian lite.
The difference between the two is that Raspbian contains a pre-installed desktop
environment, and Raspbian lite is by default only accessible through a 
command shell. 
Both Raspbian and Raspbian lite will both by default be remotely
accessible using \ac{SSH}. \ac{SSH} provides the means to remotely
connect to another machine through a command shell. In fact it is the successor
of \ac{Telnet} that has been designed with network security in mind.

Our testbed does not have monitors so we will install Raspbian lite.
A desktop environment can be installed using Raspbian's package manager
(apt-get) in case a desktop environment will be desired sometime in the future.

Below is the procedure we did to setup the \ac{Raspi}s in our testbed. We
have performed a simple setup where each \ac{Raspi} is equipped with a
memory card that contains a slightly customized Raspbian lite image with
identical settings.

@@

The overall procedure to customize the official Raspbian lite image
are:
\begin{enumerate}
    \item Download Raspbian lite
    \item Alter Raspbian lite. E.g. browse, modify, add, and delete files
        in the official Raspbian lite image
    \item Change root. I.e. change root filesystem into the Raspbian lite
        image to update and install software packages
    \item Write image to memory cards
\end{enumerate}

Hashmark (\#) and dollar (\$) will be used throughout the procedure
to indicate if an operation should be run with root permissions or
user permissions.
%\textbf{hashmark (\#) means to perform operation as root, dollar (\$) means
%to perform operation as user}

@@

After the terminal has been opened, we first declare a few variables to reduce
redundant typing. The first variables we declare is the Raspbian image name.
Notice that the extension has been omitted. This is because the image has been
packed into a zip file.
The other variable we declare is a working directory. This is where we will
download the image to and work on it. In other words, it will be stored in
/home/<username>/Raspbian

>>>>

%This section will present a testbed of thirty \ac{Raspi}s and a \ac{PC}
The testbed consists of thirty \ac{Raspi}s that will be setup and controlled
by a testbed administrator from a \ac{PC}. The \ac{Raspi}s are each equipped
with a memory card and connected to the
university network by wire using their ethernet interface (eth0). To easily
discover the \ac{Raspi}s within the university network,
they have all been assigned a unique name under a common domain.
This eliminates the need to remember their individual \ac{IP} addresses.
A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
We have two types of \ac{Raspi}s in our setup.
Table~\ref{tab:rasp_naming} shows the naming convention.

\begin{table}[ht!]
  \centering
  \begin{tabular}{|c|c|}
    \hline
      \textbf{Hostname} & \textbf{Model} \\ \hline
    rasp00 - rasp19 & Raspberry Pi 1 model B rev. 2 \\ \hline
    rasp20 - rasp29 & Rasperry Pi 2 model B V1.1 \\ \hline
  \end{tabular}
  \caption{Raspberry Pi naming convention}
  \label{tab:rasp_naming}
\end{table}

@@

The testbed should be usable by students to perform simulations and
measurements as part of their semester projects. It is therefore desired
that the \ac{Raspi}s are configured identically, but still allows the
students to alter the configurations. When working on a testbed, it is
not always desired to make persistent
changes to the devices. We also present how the concept of stacked
filesystems can be used to enable the means of both persistent and
temporary storage.

@@

To get started, we first need to install an \ac{OS} on the
\ac{Raspi}s. We will use Raspbian Linux~\cite{raspbian}. We will be
using a \ac{PC} running a Debian-based Linux distribution
to setup a customized Raspbian Linux image that can be written to
a memory card and work in any \ac{Raspi} version 1 or 2.
%We will use a desktop that has linux installed to download,
%configure, and write the image to the memory cards.
%We have decided to use a linux desktop machine to create a
%common image of Raspbian Linux that can be used in any of the Raspberry Pis.

@@

Raspbian is made available in two bundles: Raspbian and Raspbian lite.
The difference between the two is that Raspbian contains a pre-installed desktop
environment, and Raspbian lite is by default only accessible through a 
command shell. Both Raspbian and Raspbian lite will both by default be remotely
accessible using \ac{SSH}. \ac{SSH} provides the means to remotely
connect to another machine through a command shell. In fact it is the successor
of \ac{Telnet} that has been designed with network security in mind.

Our testbed does not have monitors so we will install Raspbian lite.
A desktop environment can be installed using Raspbian's package manager
(apt-get) in case a desktop environment will be desired sometime in the future.
Below is the procedure we did to setup the \ac{Raspi}s in our testbed. We
have performed a simple setup where each \ac{Raspi} is equipped with a
memory card that contains a slightly customized Raspbian lite image with
identical settings.

@@

The overall procedure to customize the official Raspbian lite image
are:
\begin{enumerate}
    \item Download Raspbian lite
    \item Alter Raspbian lite. e.g. browse, modify, add, and delete files
        in the official Raspbian lite image
    \item Change root, i.e. change root filesystem into the Raspbian lite
        image to update and install software packages
    \item Write image to memory cards
\end{enumerate}

In the following, the hashmark (\#) and dollar (\$) signs will be used
throughout the procedure to indicate if an operation should be run with
root permissions or user permissions.
%\textbf{hashmark (\#) means to perform operation as root, dollar (\$) means
%to perform operation as user}

@@

After the terminal has been opened, we first declare a few variables to reduce
redundant typing. The first variables we declare is the Raspbian image name.
Notice that the extension has been omitted. This is because the image has been
packed into a zip file. The other variable we declare is a working directory.
This is where we will download the image to and work on it. In other words, it
will be stored in \texttt{/home/<username>/Raspbian}
